---
title: "Credit card Market Segmentation by Auto-Clustering"
author: "Kar Ng"
date: '2022-05'
output: 
  github_document: 
    toc: true
    toc_depth: 4
always_allow_html: yes
---


## 1 R PACKAGES

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(skimr)
library(factoextra)
library(hopkins)
library(clValid)
library(NbClust)

```


## 2 INTRODUCTION

This project will use clustering technique to develop a customer segmentation to define marketing strategy. 

Dataset used in this project is called "Credit Card Dataset for Clustering" by Arjun Bhasin. It is a public dataset acquired from [Kaggle.com](https://www.kaggle.com/datasets/arjunbhasin2013/ccdata).

The dataset has 17 behavioral information of about 9000 active credit card holders.

## 3 DATA PREPARATION

### 3.1 Data Import

Following codes import the dataset and specify the first column as row's name, as it is required to perform clustering. 

```{r}
cc <- read.csv("cc_dataset.csv",
               header = T,
               row.names = 1)   # specificy column 1 as row name 

```

randomly sample the first 10 rows of the dataset. 

```{r}
sample_n(cc, 10)

```

The name of all variables are:

```{r}
names(cc)

```


### 3.2 Data Description

Following is the data description extracted from the kaggle website.

```{r}

Variables <- c("CUSTID", "BALANCE", "BALANCEFREQUENCY", "PURCHASES", "ONEOFFPURCHASES",
               "INSTALLMENTSPURCHASES", "CASHADVANCE", "PURCHASESFREQUENCY", "ONEOFFPURCHASESFREQUENCY", "PURCHASESINSTALLMENTSFREQUENCY", "CASHADVANCEFREQUENCY", "CASHADVANCETRX", "PURCHASESTRX", "CREDITLIMIT", "PAYMENTS", "MINIMUM_PAYMENTS", "PRCFULLPAYMENT", "TENURE")

Description <- c("Identification of Credit Card holder (Categorical)",
                 "Balance amount left in their account to make purchases",
                 "How frequently the Balance is updated, score between 0 and 1 (1 = frequently updated, 0 = not frequently updated)",
                 "Amount of purchases made from account",
                 "Maximum purchase amount done in one-go",
                 "Amount of purchase done in installment",
                 "Cash in advance given by the user",
                 "How frequently the Purchases are being made, score between 0 and 1 (1 = frequently purchased, 0 = not frequently purchased)",
                 "How frequently Purchases are happening in one-go (1 = frequently purchased, 0 = not frequently purchased)",
                 "How frequently purchases in installments are being done (1 = frequently done, 0 = not frequently done)",
                 "How frequently the cash in advance being paid",
                 "Number of Transactions made with Cash in Advanced",
                 "Number of purchase transactions made",
                 "Limit of Credit Card for user",
                 "Amount of Payment done by user",
                 "Minimum amount of payments made by user",
                 "Percent of full payment paid by user",
                 "Tenure of credit card service for user")


data.frame(Variables, Description) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("bordered", "stripped"))

```

### 3.3 Data Exploration

**Data Size and type**

The dataset contain 8950 rows and 17 variables. The "dbl" and "ind" are data type allocated by R to a particular column. The "dbl" stands for "double", it is used for numerical variables that have decimal places. The "int" stands for integer, it is used for numerical variables that have integer values. All variables are numeric with either "dbl" and "int", they will be treated the same type during analysis.

```{r}
glimpse(cc)

```
**Purchases**

During exploration, I found that "PURCHASES" is the sum of "ONEOFF_PURCHASES" and "INSTALLMENTS_PURCHASES". It may not be important in this analysis.

Following select 10 rows among the dataset, and the new variable "MY_PURCHASES" proves my finding, which is the same as the "PURCHASES", and is the sum of "ONEOFF_PURCHASES" and "INSTALLMENTS_PURCHASES".

```{r}
cc %>% 
  top_n(10, BALANCE) %>% 
  dplyr::select(ONEOFF_PURCHASES, INSTALLMENTS_PURCHASES, PURCHASES) %>% 
  mutate(MY_PURCHASES = ONEOFF_PURCHASES + INSTALLMENTS_PURCHASES)

```

**Missing values check**

Again, proven by other function that the tables have 8950 rows of data and 17 variables. All variables are numerical, and, by examining the variables "n_missing" and "complete_rate" in following tables, there is 1 missing value in the variable "CREDIT_LIMIT" and 313 in the "MINIMUM_PAYMENTS". These missing values need to be handled. 

```{r}
skim_without_charts(cc)

```
Alternatively, following code performs the missing-value check.

```{r}
colSums(is.na(cc))

```
There are 8950 rows of data and I will still have 96.5% of data left after removal of these missing values and therefore I will simply remove these missing values for the simplicity of this project.

Typically, missing value can be handled by either removal, replaced with mean, median, or using imputation algorithm such as KNN or bagging algorithm. These techniques are usually performed when there are too many missing values in important variables. For example, when missing values is higher than 5% and less than 60%. 


## 4 DATA CLEANING

### 4.1 Rename all variables

The name of all variables are in capital form and which would be difficult to read for readers and also myself.

```{r}
colnames(cc)
```
Following code transforms all the name into reading-friendly form. 

```{r}
cc <- cc %>% 
  rename_all(str_to_sentence)

```

Checking again the name of each variable.

```{r}
colnames(cc)

```

### 4.2 NA Removal

Following code remove all the missing values in the dataset (314 rows among 8950 rows)

```{r}
cc <- cc %>% 
  na.omit()

```

Now, the number of rows have been reduced to 8636 from 8950. 

```{r}
str(cc)
```

## 5 EDA

**Distribution Study**

A primary exploratory data analysis is suggested to quickly understand the general distribution of the data.  

```{r, fig.height=10, fig.width=13, message=FALSE, warning=FALSE}

# data frame

cc2 <- cc %>% 
  pivot_longer(c(1:17), 
               names_to = "my.variable", 
               values_to = "my.value")

# graphs

ggplot(cc2, aes(x = my.value, fill = my.variable)) +
  geom_histogram(color = "black") +
  facet_wrap(~my.variable, scale = "free") +
  theme_minimal() +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, face = "bold")) +
  labs(title = "Variable Distribution Analysis",
       subtitle = "by Histogram")


```

Most of the variables follow Pareto trend (80:20 rule) with majority of the data belong to one side of the value. it can be seem quite hard to group the data and categories the data into several distinct group for marketing purposes.


## 6 Clustering

The dataset must met several conditions prior to be clustered,

1. Observation as row and variable as variable, and it has been met.  
2. No missing values in the dataset, and it has been met.  
3. Standardising the data (generally recommended) to make variables comparable. This step will transform the data in all variables to a scale that would have 0 mean and 1 standard deviation. This step will be performed in this section.    

Standardising the dataset:

```{r}
cc.scale <- scale(cc)
```

Showing the first 6 rows of the data:

```{r}
head(cc.scale)

```

There are different type of clustering methods such as partitioning clustering which include "k-means clustering", "PAM", and "CLARA", or hierarchical clustering which include the famous "AGNES" and "DIANA". 

Partitioning and hierarchical clustering methods are alternative to each other, and although automation with selection statistics are available to choose the best algorithm for the dataset but I will not perform it here, because I will perform an advanced "Hybrid" clustering technique for this project, called "Hierarchical K-Means Clustering (hkmeans)". This technique combine both the algorithms.


### 6.1 Clustering Tendency Assessment

There is a big issue with clustering techniques is that it will always return clusters even if there is no any cluster in the dataset. Therefore, clustering tendency assessment is an important step to test is there clusters in the dataset. 

**Hopkins Statistics** is a statistical method that using probability to test for spatial randomness of the data. The null hypothesis is that the credit card dataset is uniformly distributed, indicating no meaningful cluster. 

* If the value of hopkins statistic is low (0 - 3), it indicates regularly-spaced data  
* if the value is around zero, it indicates random data   
* If the value is around 0.7 to 1, it indicate clustered data  

```{r}
hopkins(cc.scale, m = nrow(cc.scale)-1)

```
The Hopkins Statistics is 1, and base on this value I can reject the null hypothesis and conclude that the credit card dataset is significantly a clusterable data.

### 6.2 Find Optimal K

This section will use two typical methods among other popular methods to find the optimal number of clusters (specified by "k") that the dataset is able to be clustered into. All methods should recommend the same or highly similar value of optimal k.

In general, the optimal k is specified by the analyst of the project, and it can a hard task. There is also no definite answer to say what k is the best, it is subjective and relying on the selection of the distance matrix and the parameters used for clustering the data.


#### 6.3.1 Elbow Method

Following shows the result of Elbow method. In Elbow method, clustering algorithm is computed for different number of clusters (k). For each k, the total within sum of square (WSS) is calculated.

The location of a bend (knee) is generally considered as an indicator of appropriate k. 

```{r}

g1 <- fviz_nbclust(cc.scale, FUNcluster = kmeans, method = "wss")

g1 + geom_vline(xintercept = 7, linetype = 2)

```

The optimal number of k might be 7. 


#### 6.3.2 Silhouette Method

This method uses average silhouette width to determine the optimal number of k. A high average silhouette width means good clustering. Additionally, the algorithm will suggest the best k for the credit card dataset, which is 9, although k-7 was also not too bad in term of average silhouette width and is not too far away k-9.

```{r}
fviz_nbclust(cc.scale, 
             FUNcluster = kmeans, 
             method = "silhouette")
```


### 6.3 Hierarchical K-Means Clustering


#### 6.3.1 hkmeans Algorithms

Base on the suggestions gained from Elbow and Silhouette method, I have decided to use 7 as the optimal number of k. 

In this section, I will perform hkmeans, which is a method that combine K-means clustering and hierarchical clustering to improve k-means results. 

The biggest limitation K-means clustering is that the final result is subjected to the initial random selection of centroids. However, in hkmeans, the selection of centroid is not based on randomisation but the results of hierarchical clustering from the hk-mean algorithm, these centroids are optimiased centroids. Therefore, in hierarchical k-means clustering, the hierarchial clustering is being performed first, followed by the k-means clustering. 

Following summarise the combination of the two algorithms :

* **Step 1**: Compute hierarchical clustering with the optimal K obtained from previous section. In this step, distance matrix between each pair of observation is computed and based on that, a linkage function is selected to group similar objects into hierarchical cluster tree. Each observation are considered a cluster on its on, then the forming of clustering move upward and form bigger cluster, and this step continues until the root is reached. Finally, the tree is cut into pre-specified K (which is 7 in this case). 

* **Step 2**: Compute the center (i.e. mean) of each cluster. Therefore, there are 7 new center here in each of the 7 clusters.

* **Step 3**: Compute K-means clustering on the original normalised credit card dataset by using the set of cluster centers computed in step 2 as the initial cluster centers. Then, according to the algorithm of K-means clustering, each observations will be assigned to the closest centroid based on distance matrix (eg. euclidean). For each k-clusters, a new mean is being updated again by calculating the new mean value of all data points in the cluster. Finally, all observations are being re-allocated to these new means. The allocation of observations and the calculation of new means are iterated until the total within sum of square is minimised.

#### 6.3.2 Codes

Following code complete the hkmeans.

```{r}
res.hk <- hkmeans(cc.scale, 7)

summary(res.hk)

```





## REFERENCE

Alboukadel Kassambara 2017， *Practical Guide to Cluster Analysis in R*, Multivariate Analysis 1, Edition 1, sthda.com


