---
title: "cc_cluster"
author: "Kar Ng"
date: '2022-05'
output: 
  github_document: 
    toc: true
    toc_depth: 4
always_allow_html: yes
---


## R PACKAGES

```{r}
library(tidyverse)
library(kableExtra)
library(skimr)

```


## INTRODUCTION

This project will use clustering technique to develop a customer segmentation to define marketing strategy. 

Dataset used in this project is called "Credit Card Dataset for Clustering" by Arjun Bhasin. It is a public dataset acquired from [Kaggle.com](https://www.kaggle.com/datasets/arjunbhasin2013/ccdata).

The dataset summarises the usage behavior of about 

## DATA PREPARATION

### Data Import

Following codes import the dataset and specify the first column as row's name, as it is required to perform clustering. 

```{r}
cc <- read.csv("cc_dataset.csv",
               header = T,
               row.names = 1)   # specificy column 1 as row name 

```

randomly sample the first 10 rows of the dataset. 

```{r}
sample_n(cc, 10)

```

### Data Description



```{r}

Variables <- c("CUSTID", "BALANCE", "BALANCEFREQUENCY", "PURCHASES", "ONEOFFPURCHASES",
               "INSTALLMENTSPURCHASES", "CASHADVANCE", "PURCHASESFREQUENCY", "ONEOFFPURCHASESFREQUENCY", "PURCHASESINSTALLMENTSFREQUENCY", "CASHADVANCEFREQUENCY", "CASHADVANCETRX", "PURCHASESTRX", "CREDITLIMIT", "PAYMENTS", "MINIMUM_PAYMENTS", "PRCFULLPAYMENT", "TENURE")

Description <- c("Identification of Credit Card holder (Categorical)",
                 "Balance amount left in their account to make purchases",
                 "How frequently the Balance is updated, score between 0 and 1 (1 = frequently updated, 0 = not frequently updated)",
                 "Amount of purchases made from account",
                 "Maximum purchase amount done in one-go",
                 "Amount of purchase done in installment",
                 "Cash in advance given by the user",
                 "How frequently the Purchases are being made, score between 0 and 1 (1 = frequently purchased, 0 = not frequently purchased)",
                 "How frequently Purchases are happening in one-go (1 = frequently purchased, 0 = not frequently purchased)",
                 "How frequently purchases in installments are being done (1 = frequently done, 0 = not frequently done)",
                 "How frequently the cash in advance being paid",
                 "Number of Transactions made with Cash in Advanced",
                 "Number of purchase transactions made",
                 "Limit of Credit Card for user",
                 "Amount of Payment done by user",
                 "Minimum amount of payments made by user",
                 "Percent of full payment paid by user",
                 "Tenure of credit card service for user")


data.frame(Variables, Description) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("bordered", "stripped"))

```



### Data Exploration


## Unsupervised Clustering


